---
title: "Week 05 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exercises

```{r}
if(! require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse,
  here,
  tidycensus
)
```

1. Read COVID deaths data from the provided URL into an object called `covid`.

```{r}

url <- read.csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"))
```

```{r}
url
```


2. There are several columns that we will not need to work with or that should be renamed to improve the clarity of the fields.

Remove:

-   `UID`
-   `iso2`
-   `iso3`
-   `code3`
-   `Country_Region`
-   `Lat`
-   `Long_`
-   `Combined_Key`

Rename:

-   `Admin2` to `County`

Write this to an object called `covid`.

```{r}
#first removing specific columns and putting them into a temporary variable

urltemp <-
  select(url, -UID, -iso2, -iso3, -code3, Country_Region, -Lat, -Long_, -Combined_Key)



#renaming Admin2 to 'County' and assigning it to the variable 'covid'

covid <- rename(urltemp, County = Admin2)
  

```




3. Utilize the `tidycensus` package to read American Community Survey data at the county level for 2019 for the variable `B22007_002`. Call this new object `food_stamps`. It will be helpful downstream to rename the `estimate` and `moe` variables as `food_stamps_estimate` and `food_stamps_moe`. We also will not need the `variable` field.

```{r} 
#setting up my libraries

library(tidyverse)
library(tidycensus)

census_api_key("dc7d515284d2934987f695bf00a56908f298cbd7", overwrite = TRUE)

##I've been having trouble getting the census data installed into Rstudio, I've tried to complete this homework the best that I can from my class notes last week but I haven't been able to really run anything past this point so this is all my best guess
##

```

```{r}
get_acs(geography = "county",
        variables = "B22007_002")
```


```{r}
food_stamps <-
  load_variables(2019, 'keyfromacs')
```

```{r}
food_stamps %>% 
  filter(name == "B22007_002")
```
```{r}
food_stamps <-
  rename(food_stamps, estimate = food_stamps_estimate) %>% 
  rename(food_stamps, moe = food_stamps_moe )
```


4. Re-shape the `covid` dataframe from wide to long such that there is a `date` column and a `deaths` column. Write this to a new dataframe called `covid_long`.

```{r}
covid_long <-
  covid %>% 
  pivot_longer(names_from = variable,
              values_from = estimate)
```

5. The `FIPS` value in the `covid_long` dataset and the `GEOID` value in the `food_stamps` dataset represent the same unique ID for each county in the US. Currently, we cannot join the two fields together, though. What is the reason why this is not possible?

```{r}
##We cannot join the fields together because the FIPS field is a longer field of numbers specifying the exact area whereas the GEOID is a two digit code identifying the specific state
##
```


Adjust the `FIPS` field in the `covid_long` dataset with the following code to fix the issue. Describe what this code is doing and why this will fix the initial issue.

```{r}
covid_long <- 
  covid_long %>% 
  mutate(FIPS = str_pad(FIPS, 5, pad = "0"))
```

```{r}
##This code changes the FIPS string so that after the first five digits in the FIPS field a zero is inserted to seperate those first 5 characters with the rest of the string
##
```


6. Join the `food_stamps` dataframe to the `covid_long` dataframe. Identify the field in each dataframe that will be used as the primary key. Call this new dataframe `covid_joined`. After completing the join, you may notice the `County` and `Province_State` fields in relation to the `NAME` field as a confirmation that you successfully completed your join.

```{r}

covid_joined <-
  left_join(x = food_stamps,
          y = covid_long,
          suffix = c("_foodstamps", "_covid"))
```
7. Our new data frame will require some light cleaning as well:

-   Remove the `NAME` field
-   It may be helpful to reorder your columns so that like information is together
-   Some COVID deaths are not assigned to a county. Filter out any counties that are "Unassigned"
-   The `date` field needs to be converted to type Date. Use the following code to enable this:
    -   `date = as.Date(date, format = "%m/%d/%y")`
    -   We'll discuss dates much more in the next class.
-   Assign this back to the `covid_joined` object.

```{r}
#figure out what column names are available
colnames(covid_joined)
#create a variable to group them in an intuitive way
col_order <- c("column1a","column1b", "column2a", "column2b")


covid_joined <-
  select(-NAME) %>% #remove 'NAME' column
  covid_joined[, col_order] %>%  ##reorder the columns how you see fit
  filter(!county %in% c("Unassigned")) %>% #filter out unassigned counties
  date = as.Date(date, format = "%m/%d/%y") #fix the dates


  
```

8. How have the number of deaths changed over time for Delaware counties? Build a graph to support your answer. Interpret the graph.

```{r}
covid_joined %>% 
  select('DEcountyIDs','Date','Deathrate_county')
   %>% 
  ggplot(aes(x = Date,
             y = Deathrate/count)) +
  geom_col(fill = "DEcountyIDs") 
 
```

9. Consider counties in Delaware, Maryland, Pennsylvania, New York, and New Jersey.

What is the per capita death rate for each county? Which county has the highest per capita death rate?

```{r}
covid_joined %>% 
  mutate(county_deathrates = Deathrate_county/population_county) %>% 
  arrange(desc(county_deathrates)) %>% 
  ggplot(aes(x = county_deathrates)) +
  geom_bar() 
```

10. Create a graph that demonstrates the relationship between food stamps per capita and deaths per capita for Delaware, Maryland, Pennsylvania, New York, and New Jersey. Interpret the graph.

```{r}
covid_joined %>% 
  mutate(state_death_rates = Deathrate_state/population_state) %>% 
  mutate(state_foodstamp_rates = food_stamps_state/population_state)%>% 
  ggplot(aes(x = states,
             y = count)) +
  geom_bar(fill = (state_death_rates,state_foodstamp_rates),
           position = "avoid")
```
